---
title: "EDA"
author: "Katrina Truebebach"
date: "March 16, 2019"
output: html_document
---

```{r}
rm(list = ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(stringr)
library(lubridate)
library(modelr)
library(tis)
```

## Load cleaned data

```{r load}
load(file = '~/DS5110/data/proj_cleaned_dta.RData')
```

## Oscars 

Graph number of Oscars for actors and directors against real revenue. Boxplot and bar plot (average revenue)
_Both are linear but very weak. Unclear if should include in model_


```{r graph_oscarnum}
# Versions of data with average revenue by number of oscars 
train_oscar_actor <- train %>%
  group_by(total_oscars_actor) %>% 
  summarize(avg_real_gross = mean(real_gross))
train_oscar_director <- train %>%
  group_by(total_oscars_director) %>% 
  summarize(avg_real_gross = mean(real_gross))

# Functions to graph number of Oscars held by actors in movie vs. real revenue 
# boxplot
oscar_box <- function(df, var, title_str, x_str) {
  ggplot(df, aes_string(var, "real_gross")) + 
    geom_boxplot() + 
    labs(title = title_str, x = x_str, y = 'Real Gross Revenue ($)')
}

# bar graph 
oscar_bar <- function(df, var, title_str, x_str) {
  ggplot(df, aes_string(var, "avg_real_gross")) + 
    geom_col() + 
    labs(title = title_str, x = x_str, y = 'Average Real Gross Revenue ($)')
}

# actors
oscar_box(train, 'total_oscars_actor', 'Total # Oscars of Actors vs Real Gross Revenue', 
          'Total # Oscars Held by Actors in Movie')
oscar_bar(train_oscar_actor, 'total_oscars_actor', 'Total # Oscars of Actors vs Real Gross Revenue', 
          'Total # Oscars Held by Actors in Movie')

# directors
oscar_box(train, 'total_oscars_director', 'Total # Oscars of Directors vs Real Gross Revenue', 
          'Total # Oscars Held by Directors in Movie')
oscar_bar(train_oscar_director, 'total_oscars_director', 'Total # Oscars of Directors vs Real Gross     Revenue', 'Total # Oscars Held by Directors in Movie')

```

## Year 

Average real revenue vs year  
Adeed APPROXIMATE recession shading. Annual data, so hard to do.  
Real revenue increase during recessions (have seen this before with Great Depression - numerous articles we can reference)  
Regardless, clear that year could have an effect

```{r year}
# take average of revenue per year
train_sum <- train %>%
  group_by(year) %>%
  summarize(real_gross_avg = mean(real_gross))

# need to limit because before 1995 there are very few observations per year (< 10 usually). 
# this causes large spikes because one high earning or low earning movie influences the average heavily
# Starting at 1995, where have at least 30 (or very close) movies per year. Now can see movements over time 

ggplot(data = train_sum %>% filter(year >= 1995)) + 
  geom_rect(aes(xmin = 2008, xmax = 2010, 
                ymin = min(real_gross_avg, na.rm = T), 
                ymax = max(real_gross_avg, na.rm = T)), alpha = .05) +
  geom_rect(aes(xmin = 2001, xmax = 2002, 
                ymin = min(real_gross_avg, na.rm = T), 
                ymax = max(real_gross_avg, na.rm = T)), alpha = .05) + 
  geom_line(aes(x = year, y = real_gross_avg)) +
  labs(title = 'Average Real Gross Revenue Over Time', x = 'Year', y = 'Real Gross Revenue ($)') 

```

## Content Rating 

Bar graph of average real revenue and boxplot   
_Linear relationship. Good candidate to include in the model_

```{r content_rating}

# data manipulation. Factor. 
train_content <- train %>%
  # filter out missing
  filter(!is.na(content_rating)) %>%
  # Make content rating a factor so can order in graphics
  mutate(content_rating = as.factor(content_rating), 
         content_rating = reorder(content_rating, real_gross))

# boxplot 
train_content  %>% 
  ggplot() + 
  geom_boxplot(aes(x = content_rating, y = real_gross, fill = content_rating)) +
  labs(x = 'Content Rating', y = 'Real Gross Revenue ($)', 
       title = 'Content Rating vs Real Gross Revenue') + 
  theme(legend.position = 'none')

# bar graph
train_content  %>% 
  # average revenue by content rating
  group_by(content_rating) %>%
  summarize(avg_real_gross = mean(real_gross)) %>%
  ggplot() + 
  geom_col(aes(x = content_rating, y = avg_real_gross, fill = content_rating)) +
  labs(x = 'Content Rating', y = 'Average Real Gross Revenue ($)', 
       title = 'Content Rating vs Average Real Gross Revenue') + 
  theme(legend.position = 'none')

```
