---
title: "EDA"
author: "Katrina Truebebach"
date: "March 12, 2019"
output: html_document
---

```{r}
rm(list = ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(stringr)
library(lubridate)
library(modelr)
library(tis)
```

## Load cleaned data

```{r load}
load(file = '~/DS5110/data/proj_cleaned_dta.RData')
```

Graph number of Oscars for actors and directors against  
_Both are linear but very weak. Unclear if should include in model_


```{r graph_oscarnum}

# Function to graph number of Oscars held by actors in movie vs. real revenue 
oscar_num <- function(df, var, title_str, x_str) {
  ggplot(df, aes_string(var, "real_gross")) + 
    geom_boxplot() + 
    labs(title = title_str, x = x_str, y = 'Real Gross Revenue ($)')
}

# actors
oscar_num(train, 'total_oscars_actor', 'Total # Oscars of Actors vs Real Gross Revenue', 
          'Total # Oscars Held by Actors in Movie')
# directors
oscar_num(train, 'total_oscars_director', 'Total # Oscars of Actors vs Real Gross Revenue', 
          'Total # Oscars Held by Actors in Movie')
```

Average real revenue vs year  
Adeed APPROXIMATE recession shading. Annual data, so hard to do.  
Real revenue during recessions (have seen this before with Great Depression - numerous articles we can reference)  
Regardless, clear that year could have an effect

```{r year}
# take average of revenue per year
train_sum <- train %>%
  group_by(year) %>%
  summarize(real_gross_avg = mean(real_gross))

# need to limit because before 1995 there are very few observations per year (< 10 usually). 
# this causes large spikes because one high earning or low earning movie influences the average heavily
# Starting at 1995, where have at least 30 (or very close) movies per year. Now can see movements over time 



ggplot(data = train_sum %>% filter(year >= 1995)) + 
  geom_rect(aes(xmin = 2008, xmax = 2010, 
                ymin = min(real_gross_avg, na.rm = T), 
                ymax = max(real_gross_avg, na.rm = T)), alpha = .05) +
  geom_rect(aes(xmin = 2001, xmax = 2002, 
                ymin = min(real_gross_avg, na.rm = T), 
                ymax = max(real_gross_avg, na.rm = T)), alpha = .05) + 
  geom_line(aes(x = year, y = real_gross_avg)) +
  labs(title = 'Average Real Gross Revenue Over Time', x = 'Year', y = 'Real Gross Revenue ($)') 



```
